Music Recommendation System Using the Million Song Dataset

### 2015-05-08 writeup/MSDgroupwriteup.tex
updated PDF, some small changes

\begin{abstract}
We attempted to build a recommendation system to recommend songs to users using data from the Million Song Dataset available on Kaggle. The main focus of this paper was to use collaborative filtering algorithms that only utilize user feedback in order to predict what songs users may like. For benchmarking the algorithms, the same Mean Average Precision truncated at 500 recommended songs used for the original Kaggle competition was chosen. It was discovered that probablistic matrix factorization did not improve results much from using a baseline of simply recommending popular songs with MAP values around 0.014, while artist-based popularity along with user-based and item-based collaborative filtering methods yielded much better results exceeding a MAP value of 0.04.

@@

\subsection{Data and Statistics}
The main data source for this project is the Million Song Data Set (MSD)  \citep{Bertin-Mahieux2011}. We used the sample data set from the Kaggle competition which contains tuples of user id, song id, and play count. We also used data which links songs to artists through track ids to calculate the artist-based popularity baseline MAP score. The links to these datasets can be found in the Appendix.

@@

\begin{table}[h]
\begin{center}
\begin{tabular}{lll}
\hline
Measurement & # of Listeners Per Song & # of Songs Played Per User \\
\hline
Mean  &     8.890194 & 13.190300  \\
Std      &    46.780258 & 8.070827 \\
Min      &    1.0  & 5.0  \\
Max     &   5043 &  53 \\
\end{tabular}
\caption{Basic summary statistics for the full data set.}\label{tab:stats}
\end{center}
\end{table}

The scale of the data was too large to explore on our computers for the algorithms that we used, so we subset the data down to a size easily handled by an average laptop.

It can be seen in Figure \ref{fig:csongs} that the majority of the songs are listened to by very few users so we can decrease the number of songs in our subset without losing all of the play counts.

>>>>

\begin{abstract}
We attempted to build a recommendation system to recommend songs to users using data from the Million Song Dataset available on Kaggle. The main focus of this paper was to use collaborative filtering algorithms that only utilize user feedback in order to predict what songs users may like. For benchmarking the algorithms, the same Mean Average Precision truncated at 500 recommended songs used for the original Kaggle competition was chosen. It was discovered that probabilistic matrix factorization did not improve results much from using a baseline of simply recommending popular songs with MAP values around 0.014, while artist-based popularity along with user-based and item-based collaborative filtering methods yielded much better results exceeding a MAP value of 0.04.

@@

\subsection{Data and Statistics}
The main data source for this project is the Million Song Data Set (MSD)  \citep{Bertin-Mahieux2011}. We used the sample data set from the Kaggle competition which contains tuples of user id, song id, and play count. We also used data which links songs to artists through track ids to calculate the artist-based popularity baseline MAP score. The links to all these datasets can be found in the Appendix.

@@

\begin{table}[h]
\begin{center}
\begin{tabular}{lll}
\hline
Measurement & \# of Listeners Per Song & \# of Songs Played Per User \\
\hline
Mean  &     8.890194 & 13.190300  \\
Std      &    46.780258 & 8.070827 \\
Min      &    1.0  & 5.0  \\
Max     &   5043 &  53 \\
\end{tabular}
\caption{Basic summary statistics for the full data set.}\label{tab:stats}
\end{center}
\end{table}

The scale of the data was too large to explore on our computers for the algorithms that we used, so we subset the data down to a size easily handled by an average laptop. 
It can be seen in Figure \ref{fig:csongs} that the majority of the songs are listened to by very few users we were able to decrease the number of songs in our subset without losing all of the play counts.
